Visible Minorities and Economic Inequality: An Examination of Rosedale-Moore Park and Downtown Yonge East

### 2024-09-26 outputs/paper/paper.qmd
working on adding figure for income vs. education

>>>>

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-five
#| fig-cap: This scatterplot depicts the relationship between the highest level of education and income categories in  Rosedale-Moore Park and Downtown Yonge East neighborhoods.
#| fig-width: 15

# install libraries 
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("ggplot2")
# install.packages("tidyr")
# install.packages("stringr")

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)

if (file.exists(file_path)) {
  analysis_data <- read_csv(file_path)
  print("File found and loaded successfully!")
} else {
  stop("Error: File not found at the specified path.")
}

# Define income and education categories
income_categories <- c("Under $10,000", "$10,000 to $19,999", "$20,000 to $29,999", 
                       "$30,000 to $39,999", "$40,000 to $49,999", 
                       "$50,000 to $59,999", "$60,000 to $69,999", 
                       "$70,000 to $79,999", "$80,000 to $89,999", 
                       "$90,000 to $99,999", "$100,000 to $149,999", 
                       "$150,000 and over")

education_categories <- c("No certificate, diploma or degree",
                          "High school diploma or equivalency certificate", 
                          "Postsecondary certificate or diploma",
                          "Apprenticeship or trades certificate or diploma",
                          "Non-apprenticeship trades certificate or diploma",
                          "College, CEGEP or other non-university certificate or diploma",
                          "University certificate or diploma below bachelor level",
                          "Bachelor's degree",
                          "University certificate or diploma above bachelor level",
                          "Degree in medicine, dentistry, veterinary medicine or optometry",
                          "Master's degree", 
                          "Earned doctorate")

# Prepare the income data
income_data <- analysis_data %>%
  filter(`Neighbourhood Name` %in% income_categories) %>%
  select(`Neighbourhood Name`, `Rosedale-Moore Park`, `Downtown Yonge East`)

# Prepare the education data
education_data <- analysis_data %>%
  filter(`Neighbourhood Name` %in% education_categories) %>%
  select(`Neighbourhood Name`, `Rosedale-Moore Park`, `Downtown Yonge East`)

# Combine income and education into a new data frame
combined_data <- data.frame(
  Education = rep(education_categories, each = length(income_categories)),  # Repeat education levels
  Income = rep(income_categories, times = length(education_categories))  # Repeat income categories
)

# Ensure the Income column is numeric by mapping the character values to their numeric indices
combined_data$IncomeNumeric <- as.numeric(factor(combined_data$Income, levels = income_categories))

# Set the Education variable as a factor to maintain order
combined_data$Education <- factor(combined_data$Education, levels = education_categories)

# Create the line plot
line_plot <- ggplot(combined_data, aes(x = Education, y = IncomeNumeric)) +
  geom_line(aes(group = 1), color = "blue") +  # Line connecting the points
  geom_point(size = 2, color = "red") +  # Points to mark each category
  scale_y_continuous(breaks = seq_along(income_categories), 
                     labels = income_categories) +  # Set y-axis labels
  labs(
    title = "Line Plot of Income Across Education Levels",
    x = "Education Levels",
    y = "Income Categories"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability

# Print the line plot
print(line_plot)

#trying to fix issue of dataset values incorrectly being read?

# Check the income_data
print(income_data)

# Check the education_data
print(education_data)

# Check the combined_data
print(combined_data)

# they all seem to contain the expected values

# Check the structure of each data frame
str(income_data)
str(education_data)
str(combined_data)
